---
title: "ReproduceableResearch2"
author: "Markus"
date: "16 April 2017"
output: html_document
---

# Analysis of Storm Damage is the USA

# Data Processing
**The data is downloaded filtered to those incidents, where damage was registrated**
```{r echo=TRUE}
library(dplyr)
library(readr)
library(ggplot2)

fileName <- 'repdata-data-StormData.csv.bz2'
fileURL <- 'https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2FStormData.csv.bz2'

if(!file.exists(fileName)) {
    download.file(fileURL, fileName, method = 'curl')
}

stormData <- read.csv(bzfile(fileName),stringsAsFactors = FALSE)

stormData <- stormData %>%
    filter(FATALITIES > 0 | INJURIES > 0 | PROPDMG > 0 | CROPDMG > 0)
```

# Results
The data is filtered by Fatalities
```{r echo=TRUE}
storm_damage <- stormData %>%
    group_by(EVTYPE) %>%
    summarise(FATALITIES = sum(FATALITIES), INJURIES = sum(INJURIES), PROPDMG = sum(PROPDMG),
              CROPDMG = sum(CROPDMG))

storm_damage <- storm_damage[order(-storm_damage$FATALITIES), ]

head(storm_damage)
```


**The data is filtered by Fatalities**
```{r echo=TRUE}
storm_damage <- stormData %>%
    group_by(EVTYPE) %>%
    summarise(FATALITIES = sum(FATALITIES), INJURIES = sum(INJURIES), PROPDMG = sum(PROPDMG),
              CROPDMG = sum(CROPDMG))

storm_damage <- storm_damage[order(-storm_damage$FATALITIES), ]
storm_damagetop5 <- storm_damage[1:5, ]

ggplot(data = storm_damagetop5, aes(x = EVTYPE, y = FATALITIES)) + geom_bar(stat="identity") + 
    ggtitle("Top 5 Wheather Events regarding Fatalities")
```


**The data is filtered by Economic Damage, which is a new variable created of Property Damage and Crop Damage**
```{r echo=TRUE}
storm_damage <- stormData %>%
    group_by(EVTYPE) %>%
    summarise(FATALITIES = sum(FATALITIES), INJURIES = sum(INJURIES), PROPDMG = sum(PROPDMG),
              CROPDMG = sum(CROPDMG))

storm_damage$ECDAMAGE <- storm_damage$PROPDMG + storm_damage$CROPDMG
storm_damage <- storm_damage[order(-storm_damage$ECDAMAGE), ]

storm_damagetop5 <- storm_damage[1:5, ]

ggplot(data = storm_damagetop5, aes(x = EVTYPE, y = ECDAMAGE)) + geom_bar(stat="identity") + 
    ggtitle("Top 5 Wheather Events regarding Economic Damage")
```

# Conclusion
* Tornados do by far the greatest damage to population
* Tornados are also doing the greatest economic damage, followed by flash floods.